---
title       : IS YOUR FOOD SAFE?
subtitle    : Outcome of Domestic Food Facilities Inspections
author      : Yi Liu
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [bootstrap]  # mathjax, quiz, bootstrap
ext_widgets : {rCharts: libraries/nvd3} # {rCharts: ["libraries/highcharts",                                              # "libraries/nvd3", "libraries/morris"]}
mode        : standalone        # {standalone, selfcontained , draft}
knit        : slidify::knit2slides

---
## Some may think food safety is not a problem in US...  
<br>

### Over **300,000** Americans are hospitalized every year due to contaminated foods or beverages  
### **5,000** out of those end up with **_death_**  
<br>
### Europe outranks U.S. on food traceability and many other aspects
e.g. In the journal _Comprehensive Reviews in Food Science and Food Safety_, food traceability is ranked as **Superior** in European Union but just **Average** in U.S.

---
## Inspections authorites have limited resourses. 

### There are **154,000** registered domestic facilities that manufacture, process, pack, or hold foods consumed by humans or animals in the United States.
<br>
### On another hand, the CFSAN is responsible for carrying out the inspections and specific tests with only a little over **800** employees. 
> Center for Food Safety & Applied Nutrition (CFSAN) is one of six centers within the Food and Drug Administration (FDA). Its major responsibility is ensuring that food is safe, nutritious and wholesome, expecially milk/dairy products.

<br>
### It is necessary to more efficiently allocate resources and detect more rapidly faulty food establishments.

---
## Counts of CFSAN inspections from 2008 to 2014
###### Certain states with abnormally good , only a few or censered ratings have been excluded.

```{r  echo=F, results='asis', fig.align='center'}
suppressPackageStartupMessages(library(googleVis))
count = aggregate(Count ~ State,readRDS('count.rds'), sum)
gv = gvisGeoChart(count, locationvar = "State",  colorvar="Count",
                options=list(region="US", 
                             displayMode="regions", 
                             resolution="provinces",
                             width=850, height=500))
print(gv, tag = 'chart')
```

---
## Possible inspection outcomes: 
<br>
###    1. Official Action Indicated (**OAI**):  
Significant objectionable conditions are found and regulatory action is warranted.
<br>
###    2. Voluntary Action Indicated (**VAI**):  
Objectionable conditions are found, but not at a level that warrants regulatory action.
<br>
###    3. No Action Indicated (**NAI**):  
No objectionable condition is found.  


---

## Will the situation be different by States?   

[State Effects](https://chaorenliuyi.shinyapps.io/Cpresnetation)
<br>
<br>
## It looks like TX did a better job than CA.
<br>
### 1. TX received a lot more NAI inspections and its NAI ratio is also higher than that in CA.
### 2. Where more inspections have been conducted, the overall rating seems to be better.
### 3. There are some clustering effects in inspections result.

---
## Yearly inspections number peaked around 2011

```{r message = F, echo = F, fig.align='center', fig.width=8,fig.height=4.5,dpi=300,out.width="1920px",height="1080px"}
library(ggplot2)
library(reshape2)
library(scales)

time.dy = readRDS('time.rds')

year.long = aggregate(X ~ Year+ District.Decision, time.dy, FUN = length )
year.count = dcast(year.long, Year~District.Decision, sum)
year.total = apply(year.count[,2:4], 1, sum)
year.ratio = as.data.frame(apply(year.count, 2, FUN = function(x){x/year.total}))
year.ratio[,1] = year.count[,1]

year.count[c(1,7), 2:4] = year.count[c(1,7), 2:4]*4

g = ggplot( year.count )
g = g + geom_line(aes(Year, NAI, colour='NAI'), size =1)
g = g + geom_line(aes(Year, VAI, colour='VAI'), size =1)
g = g + geom_line(aes(Year, OAI, colour='OAI'), size =1)
g = g + scale_colour_manual("Data Type", 
                            breaks = c('NAI', 'VAI', 'OAI'),
                            values = c('#50bd27', '#ff0241', '#be92e7'))
g = g + labs(title='CFSAN Inspection Results 2008-2014', x = "Year", y = "Percentage")
g = g + theme(plot.title = element_text(size = rel(1.5)),
              axis.title = element_text(size = rel(1.2)),
              axis.ticks = element_line(size = rel(1.2)),
              axis.text.x = element_text(size=rel(1.2)),
              axis.text.y = element_text(size=rel(1.5)),
              panel.background = element_blank())
g

```

---
## More NAI ratings were given after 2010

```{r message = F, echo = F, fig.align='center', fig.width=8,fig.height=4.5,dpi=300,out.width="1920px",height="1080px"}
g = ggplot( year.ratio )
g = g + geom_line(aes(Year, NAI, colour='NAI'), size =1)
g = g + geom_line(aes(Year, VAI, colour='VAI'), size =1)
g = g + geom_line(aes(Year, OAI, colour='OAI'), size =1)
g = g + scale_colour_manual("Data Type", 
                            breaks = c('NAI', 'VAI', 'OAI'),
                            values = c('#50bd27', '#ff0241', '#be92e7'))
g = g + scale_y_continuous(labels = percent)
g = g + labs(title='CFSAN Inspection Results 2008-2014', x = "Year", y = "Percentage")
g = g + theme(plot.title = element_text(size = rel(1.5)),
              axis.title = element_text(size = rel(1.2)),
              axis.ticks = element_line(size = rel(1.2)),
              axis.text.x = element_text(size=rel(1.2)),
              axis.text.y = element_text(size=rel(1.5)),
              panel.background = element_blank())
g

```

---
## Things went wrong in CA...

```{r message = F, echo = F, fig.align='center', fig.width=8,fig.height=4.5,dpi=300,out.width="1920px",height="1080px"}
time.CA = subset(time.dy, state == "CA") 
year.long = aggregate(X ~ Year+ District.Decision, time.CA, FUN = length )
year.count = dcast(year.long, Year~District.Decision, sum)
year.total = apply(year.count[,2:4], 1, sum)
year.ratio = as.data.frame(apply(year.count, 2, FUN = function(x){x/year.total}))
year.ratio[,1] = year.count[,1]
g = ggplot( year.ratio )
g = g + geom_line(aes(Year, NAI, colour='NAI'), size =1)
g = g + geom_line(aes(Year, VAI, colour='VAI'), size =1)
g = g + geom_line(aes(Year, OAI, colour='OAI'), size =1)
g = g + scale_colour_manual("Data Type", 
                            breaks = c('NAI', 'VAI', 'OAI'),
                            values = c('#50bd27', '#ff0241', '#be92e7'))
g = g + scale_y_continuous(labels = percent)
g = g + labs(title='CFSAN Inspection Results in CA', x = "Year", y = "Percentage")
g = g + theme(plot.title = element_text(size = rel(1.5)),
              axis.title = element_text(size = rel(1.2)),
              axis.ticks = element_line(size = rel(1.2)),
              axis.text.x = element_text(size=rel(1.2)),
              axis.text.y = element_text(size=rel(1.5)),
              panel.background = element_blank())
g

```

---
## Rural beats urban in overall ratings

```{r setup,cache = F, message = F, echo = F}
require(rCharts)
knitr::opts_chunk$set(comment = NA, results = 'asis', tidy = F, message = T)
```
```{r, echo = FALSE, results = 'hide', message=FALSE, warning=FALSE}
require(slidifyLibraries)
require(rCharts)
urban = readRDS('urban.rds')
```
```{r nvd3plot, results = 'asis', comment = NA, message = F, echo = F}
n1 <- nPlot(X ~ District.Decision, group = 'urban',
            type = 'multiBarChart', data = urban)
n1$yAxis(tickFormat = "#!
        function(d) {return d3.format('.1%')(d)}
        !#")
options(RCHART_WIDTH = 800, RCHART_HEIGHT = 550)
# n1$print("chart1", include_assets=TRUE, cdn=TRUE)
n1$show('inline', include_assets = TRUE, cdn = TRUE)
```

---
## Suggestions
<br>
### Transform some resources form **TX to CA**.

<br>
### Transform some resources form **rural areas to urban areas**.

<br>
### Investigate the cause for the **clustering effects**.

---
<br>
<br>
<br>
<br>
<br>
## Thank you!

<br>
<br>

## Any questions will be welcomed!







